Make a Map for our Paper

### 2024-03-14 001-map-of-samples.Rmd
Number the Rmds, (but not their output dirs)
Update the 001-map for the new collections

```{r}
# inputs:
input_list <- list(
)
# outputs:
output_list <- list(
  map = "results/map_of_samples/map.pdf",
  cropmap = "results/map_of_samples/map-crop.pdf",
  texmap = "tex/images/map-crop.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

```{r}
nat.earth <- rast("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") #
state_prov <- st_read("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")
coastline <- st_read("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")
all_rivers <- st_read("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp") %>%
  st_zm() %>%
  st_transform(., st_crs(state_prov))
```

@@

our_rivers <- all_rivers %>%
  filter(gnis_name %in% c(
    "Sacramento River",
    "San Joaquin River",
    "Feather River",
    "Russian River",
    "Trinity River",
    "Klamath River"
    #"Deer Creek",
    #"Battle Creek"
  ) | 
    gnis_id %in% c(
      "00220293", "00237425", "00263498", "00266522",  # these are butte creek
      "00233775", "01655075"    # these are deer
     # "00218740", "00229640", "00234966"   # these are battle creek
    ) 
  ) %>%
  bind_rows(mill_candi)

@@

g <- ggplot() + 
  geom_sf(data = our_rivers, aes(colour = gnis_name)) +
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)

@@

## Crop
```{r}
# important to put them in this order and named like this
domain <- c(
  xmin = -127.9,
  xmax = -117.4,
  ymin = 36,
  ymax = 42.6
)
sf_use_s2(FALSE)
nat_crop <- crop(nat.earth, y = ext(domain))
state_subset <- st_crop(state_prov, domain)
coastline_cropped <- st_crop(coastline, domain)

@@

mm <- base_map + 
  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", size = 0.4) +
  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
  geom_label(
    data = labels_1line,
    aes(
      x = label_long + bit + (1 - hjust * 2) * kick * ( 0.65 + rank - 1),
      y = label_lat, 
      label = map_text,
      hjust = hjust
    ),
    size = 2.2,
    label.padding = unit(0.09, "lines"),
    lineheight = 0.85
  ) +
  scale_fill_manual(values = run_time_colors) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", linewidth = 1),
    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8, family = "serif"),
    axis.title.y = element_text(family = "serif", size = 10),
    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
    legend.position = "none"
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  coord_sf(
    expand = FALSE,
  ) +
  guides(fill = guide_legend(title = "Run Timing:", nrow = 2)) +
  ggspatial::annotation_north_arrow(location = "br", height = unit(10, "mm"), width = unit(5, "mm"), style = north_arrow_fancy_orienteering()) +
  ggspatial::annotation_scale(location = "br", height = unit(0.1, "cm"))

@@

inset_world <- ggplot() +
  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", size = 0.1) +
  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", size = 0.1) +
  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3, linewidth = 0.1) +
  coord_map("ortho", orientation = c(30, -114, 0), xlim = c(-150, -75), ylim = c(28, 60)) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, -1, -1), "mm")
  )
inset_world
```

@@

```{r cowplot-to-make-inset}
final_map <- ggdraw() +
  draw_plot(mm) +
  draw_plot(inset_world, x = 0.181, y = 0.139, width = 0.20, height = 0.17)

@@

file.rename(from = output_list$cropmap, to = output_list$texmap)

>>>>

```{r}
# inputs:
input_list <- list(
  oregon_rivers = "inputs/Rivers_OR/Rivers_OR.shp"
)
# outputs:
output_list <- list(
  map = "results/map_of_samples/map.pdf",
  cropmap = "results/map_of_samples/map-crop.pdf",
  texmap = "tex/images/map-crop.pdf"
)
# you can create the necessary output directories like this:
dump <- lapply(output_list, function(x)
  dir.create(dirname(x), recursive = TRUE, showWarnings = FALSE)
)
```

@@

```{r}
nat.earth <- rast("geo-spatial/HYP_HR_SR_OB_DR/HYP_HR_SR_OB_DR.tif") #
state_prov <- st_read("geo-spatial/ne_10m_admin_1_states_provinces_lines/ne_10m_admin_1_states_provinces_lines.shp")
coastline <- st_read("geo-spatial/ne_10m_coastline/ne_10m_coastline.shp")
all_rivers <- st_read("geo-spatial/NHD_Major_Rivers_and_Creeks/Major_Rivers_and_Creeks.shp") %>%
  st_zm() %>%
  st_transform(., st_crs(state_prov))

or_rivers <- st_read(input_list$oregon_rivers)
```

## Crop the basemap, state lines coast lines

```{r}
# important to put them in this order and named like this
domain <- c(
  xmin = -127.9,
  xmax = -117.4,
  ymin = 36,
  ymax = 43.25
)
sf_use_s2(FALSE)
nat_crop <- crop(nat.earth, y = ext(domain))
state_subset <- st_crop(state_prov, domain)
coastline_cropped <- st_crop(coastline, domain)

@@

# same with Blue Creek
blue_area <- c(
  xmin = -124,
  xmax = -123.5,
  ymin = 41.2,
  ymax = 41.8
)

blue_candi <- all_rivers %>%
  filter(str_detect(gnis_name, "Blue Creek")) %>%
  st_crop(blue_area)

# rogue river:
rogue <- or_rivers %>%
  filter(name == "Rogue") %>%
  mutate(gnis_name = "Rogue River")

our_rivers <- all_rivers %>%
  filter(gnis_name %in% c(
    "Sacramento River",
    "San Joaquin River",
    "Feather River",
    "Russian River",
    "Trinity River",
    "Klamath River",
    "Eel River",
    "Smith River"
    #"Deer Creek",
    #"Battle Creek"
  ) | 
    gnis_id %in% c(
      "00220293", "00237425", "00263498", "00266522",  # these are butte creek
      "00233775", "01655075"    # these are deer
     # "00218740", "00229640", "00234966"   # these are battle creek
    ) 
  ) %>%
  bind_rows(mill_candi) %>%
  bind_rows(blue_candi) %>%
  bind_rows(rogue)

@@

g <- ggplot() + 
  geom_sf(data = our_rivers, aes(colour = gnis_name)) +
  geom_sf(data = coastline_cropped) +
  geom_point(data = labels, aes(x = label_long, y = label_lat), colour = "red") +
  geom_point(data = labels, aes(x = tip_long, y = tip_lat), colour = "blue", size = 0.2)

@@

@@

mm <- base_map + 
  geom_segment(data = labels, aes(x = tip_long, y = tip_lat, xend = label_long, yend = label_lat), colour = "black", linewidth = 0.4) +
  geom_point(data = labels, aes(x = label_long + bit + (1 - hjust * 2) * kick * (rank - 1), y = label_lat, fill = run_timing), shape = 21, size = 3.5) +
  geom_label(
    data = labels_1line,
    aes(
      x = label_long + bit + (1 - hjust * 2) * kick * ( 0.65 + rank - 1),
      y = label_lat, 
      label = map_text,
      hjust = hjust
    ),
    size = 2.2,
    label.padding = unit(0.09, "lines"),
    lineheight = 0.85
  ) +
  scale_fill_manual(values = run_time_colors) +
  theme_bw() +
  theme(
    panel.border = element_rect(colour = "black", linewidth = 1),
    axis.text.x = element_text(size = 8, family = "serif", angle = 35, hjust = 1),
    axis.text.y = element_text(size = 8, family = "serif"),
    axis.title.y = element_text(family = "serif", size = 10),
    axis.title.x = element_text(family = "serif", vjust = 2, size = 10),
    plot.margin = margin(0, 0.1, 0, 0.15, "cm"),
    legend.position = "none"
  ) +
  xlab("Longitude") +
  ylab("Latitude") +
  coord_sf(
    expand = FALSE,
  ) +
  guides(fill = guide_legend(title = "Run Timing:", nrow = 2)) +
  ggspatial::annotation_north_arrow(location = "br", height = unit(10, "mm"), width = unit(5, "mm"), style = north_arrow_fancy_orienteering()) +
  ggspatial::annotation_scale(location = "br", height = unit(0.1, "cm"))

@@

inset_world <- ggplot() +
  geom_polygon(data = wrld, aes(x = long, y = lat, group = group), colour = "black", fill = "gray90", linewidth = 0.1) +
  geom_path(data = states, aes(x = long, y = lat, group = group), colour = "black", linewidth = 0.1) +
  geom_polygon(data = domain_df, mapping = aes(x = long, y = lat), colour = "red", fill = "red", alpha = 0.3, linewidth = 0.1) +
  coord_map("ortho", orientation = c(30, -114, 0), xlim = c(-150, -75), ylim = c(28, 60)) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.margin = unit(c(0, 0, -1, -1), "mm")
  )
inset_world
```

@@

```{r cowplot-to-make-inset}
final_map <- ggdraw() +
  draw_plot(mm) +
  draw_plot(inset_world, x = 0.215, y = 0.139, width = 0.20, height = 0.17)

@@

file.copy(from = output_list$cropmap, to = output_list$texmap)
